\documentclass[12pt]{exam}

\include{mypreamble}

\begin{document}

\title{Problem Set 1}
\date{Due: January 30th, 5pm}
\maketitle


\begin{questions}

\question \textbf{Primes in Arithmetic Progressions.}

Let $a,d \in \ZZ_{>0}$, and consider the arithmetic progression
\[
a,\, a+d,\, a+2d,\, a+3d,\, \dots
\]

Using computational experimentation, investigate for which pairs $(a,d)$ the arithmetic progression $a+nd$ contains infinitely many prime numbers. Try at least $6$ different pairs $(a,d)$. 

\emph{You are strongly encouraged to use SageMath.} In SageMath, the expression \texttt{i in Primes()} returns \texttt{True} if $i$ is prime and \texttt{False} otherwise.

Based on your experiments, state a conjecture characterizing all pairs $(a,d)$ for which the arithmetic progression contains infinitely many primes. Provide clear computational evidence supporting your conjecture (for example, tables, plots, or explicit data).

\emph{You are not expected to prove your conjecture.}

\question \textbf{The Well-Ordering Principle.}

Let $S \subseteq \RR$. The \emph{Well-Ordering Principle} for $S$ is the statement:

\begin{quote}
Every nonempty subset of $S$ has a least element (with respect to the usual order on $\RR$).
\end{quote}

 Give an explicit example of a nonempty subset $S \subseteq \RR$ for which the Well-Ordering Principle does not hold, and explain carefully why it fails.

\question \textbf{Divisibility Statements.}

Let $a,b,c$ be integers. Determine whether each statement is true or false. If true, give a proof. If false, give a counterexample.

\begin{parts}
\part If $a \mid c$ and $b \mid c$, then $a+b \mid c$.
\part If $a \mid b$ and $a \mid c$, then $a \mid (b+c)$.
\end{parts}

\break 

\question \textbf{Linear Diophantine Equations.}

Find \emph{all} integer solutions to the equation
\[
252x + 198y = 18.
\]

\question \textbf{Congruences.}

Let $a,b \in \ZZ$ and let $n \in \ZZ_{>0}$. We say that $a \equiv b \pmod n$ or that $a$ is \emph{congruent to $b$ modulo $n$} if $n \mid (b-a)$. Using this definition, prove the following properties. 

\begin{parts}

\part Find $a \in \{0,1, \ldots, 11\}$ such that $7^5 + a \equiv 2 \pmod{12}$.

\part Using the definition, prove the following properties:
\begin{enumerate}[(i)]
\item $a \equiv a \pmod{n}$
\item If $a \equiv b \pmod{n}$, then $b \equiv a \pmod{n}$
\item If $a \equiv b \pmod{n}$ and $b \equiv c \pmod{n}$, then $a \equiv c \pmod{n}$
\item If $a \equiv b \pmod{n}$, then $a+c \equiv b+c \pmod{n}$ for all $c \in \ZZ$.
\end{enumerate}
\end{parts}

\question \textbf{The Euclidean Algorithm.}

Let $a$ and $b$ be positive integers with $a>b$, and let
\[
b = r_0, r_1, r_2, \dots
\]
be the successive remainders in the Euclidean algorithm applied to $a$ and $b$.

\begin{parts}
\part Show that after every two steps, the remainder is reduced by at least one half. In other words, verify that
\[
r_{i+2} < \frac{1}{2} r_i
\]
for every $i=0,1,2,\dots$.

\part Conclude that the Euclidean algorithm terminates in at most $2\log_2(b)$ steps, where $\log_2$ denotes the logarithm to the base $2$.

\part Deduce that the number of steps is at most seven times the number of decimal digits of $b$.

\emph{Hint: What is the value of $\log_2(10)$?}
\end{parts}

\question \textbf{The $3n+1$ Algorithm.}

The \emph{$3n+1$ algorithm} works as follows. Start with any positive integer $n$.
\begin{itemize}
\item If $n$ is even, divide it by $2$.
\item If $n$ is odd, replace it with $3n+1$.
\end{itemize}
Repeat this process indefinitely.

For example, starting with $5$ we obtain
\[
5,16,8,4,2,1,4,2,1,4,2,1,\dots
\]
and starting with $7$ we obtain
\[
7,22,11,34,17,52,26,13,40,20,10,5,16,8,4,2,1,4,2,1,\dots
\]

Notice that if the algorithm ever reaches $1$, the sequence repeats $4,2,1$ forever.

In general, one of the following two possibilities occurs:
\begin{enumerate}
\item The sequence eventually repeats a number that appeared earlier. In this case, the block of numbers between the two occurrences repeats indefinitely. We say that the algorithm \emph{terminates} at the last nonrepeated value, and the number of distinct entries is called the \emph{length} of the algorithm.
\item The sequence never repeats any value, in which case we say that the algorithm does not terminate.
\end{enumerate}

\begin{parts}
\part Verify that the algorithm terminates at $1$ for the starting values $5$ and $7$, and compute the length of the algorithm in each case.
\part Write a short program or use computational software to test the algorithm for many starting values. What do you observe?
\part State a conjecture about whether the $3n+1$ algorithm terminates for all positive integers.
\end{parts}

\end{questions}

\end{document}
