\documentclass[12pt]{exam}

\newcommand{\spc}[1]{{\color{teal}#1}}
\include{mypreamble}

\begin{document}

\title{Problem bank}
\maketitle

\section{Simple computational questions}

\begin{question}
    \question Find all solutions to the congruence equation $5x^2 \equiv 1 \mod 11$ with $x \in \mathbb{Z}_{11} = \{0,1\ldots, 10\}$. 
\end{question}

\section{Exploratory worksheet}
\spc{I gave this in class on January 29th.}

For each of the statements below, determine whether they are true or false: give a proof if they are true or a counterexample if false. \emph{(Warning: Some statements might be difficult to prove/disprove, in which case you should guess whether you think the result is true or false and provide some computational evidence.)} 

\begin{questions}
\question Every prime number $p > 3$ is of the form $6k + 1$ or $6k-1$ for some integer $k$. 

\question Let $n > 2$ be an integer. 
\begin{parts}
    \part If $n \mid 2^{n-1} - 1$ then $n$ is prime. 
    \part If $n$ is prime, then $n\mid 2^{n-1} -1$. 
\end{parts}

\question 
\begin{parts}
    \part There are infinitely many primes of the form $3k + 1$. 
    \part There are infinitely many primes of the form $3k + 2$. 
\end{parts}

\question There are more primes of the form $3k + 2$ than of the form $3k +1$. More specifically, if 
    \[ \pi_1(n) = \#\{\text{primes $p \leq n$ of the for $3k+1$} \}, \]
    \[ \pi_2(n) = \#\{\text{primes $p \leq n$ of the for $3k+2$} \}, \]
then $\pi_1(n) \leq \pi_2(n)$ for any integer $n \geq 2$. 
\end{questions}

\section{Challenge problems}

\begin{questions}
   \question For every integer $n \geq 2$ there is a prime number between $n$ and $2n$. 

    \spc{This is Bertrands postulate.}

    \question $n$ is prime if and only if $a^n \equiv a \mod n$ for every integer $a$. 

    \spc{Fermat's little theorem, and the converse is false: Introduce Carmichael numbers. }

    \question There are infinitely many Carmichael numbers. 

    \spc{This is hard! Proven by Pomerance,...}

    \question Show that $n$ is a Carmichael number if and only if for every prime $p$ that divides $n$, $p-1$ also divides $n-1$.

    \spc{This is Korselt's criterion. I have no clue how easy it is to prove.}

    \question There are infinitely many Carmichael numbers in an arithmetic progression. 

    \spc{Also hard! Daniel Larsen's theorem. }
    
 
\end{questions}
\end{document}